---
title: "class10"
output: html_document
---

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
head(wisc.df)
```


```{r}
# Convert the features of the data: wisc.data
wisc.data <- as.matrix( wisc.df[,3:32] )
head(wisc.data)
```

> Q. How many cancer and non-cancer? 

```{r}
table(wisc.df$diagnosis)
```

> Q. How many variables/features in the data are suffixed with _mean?

For this use the grep function

```{r}
colnames(wisc.data)
grep("_mean", colnames(wisc.data), value=TRUE)
length(grep("_mean", colnames(wisc.data), value=TRUE))
```



## Principal Component Analysis

We need to scale our input before performing PCA

```{r}
round(apply(wisc.df, 2, sd),2)
```

need to set scale to true

```{r}
wisc.pr <- prcomp(wisc.data, scale = TRUE)
summary(wisc.pr)
```

> Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

The first PC captures 44.27% of the original variance. Note that 72.6% is captured in the first 3 PC's

```{r}
biplot(wisc.pr)
```


```{r}
attributes(wisc.pr)
```


```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
abline(h=0, col="gray", lty=2)
abline(v=0, col="gray", lty=2)

```


cluster in PC space

```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:3]), method="ward.D2")
plot(wisc.pr.hclust)
```



```{r}
grps <- cutree(wisc.pr.hclust,k=3)
table(grps)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=grps)
```


```{r}
table(grps, wisc.df$diagnosis)
```


## Prediction
```{r}
new <- read.csv("new_samples.csv")
```

```{r}
npc <- predict(wisc.pr,newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], labels = c(1,2), col="white")
```


































